The impact of forest management on the temperature sensitivity of organic matter turnover

### 2024-10-24 Manuscript_draft/Draft.Rmd
Minor edits on the text

Here, we hypothesized that logging of trees will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated. In turn, addition of slash will provide labile organic matter to decompose increasing the respiration rates. We also hypothesized that the variation in soil respiration would be mainly dependent on microclimatic variables (soil temperature and moisture), especially in the management treatments where vegetation was removed, since heterotrophic respiration will prevail over autotrophic respiration. Moreover, the sensitivity of soil respiration to microclimatic conditions will depend on the type of climate/forest, being the Mediterranean the most sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.

To test these hypotheses, we have monthly recorded data on soil respiration and microclimate in three replicated Quercus forests across Europe subjected to different forest management strategies during three years. We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic model…

@@

## Validation
In order to ensure a robust validation, all models have been validated with a split approach, keeping two completely independent data sets (80% of the points for training and 20% validation). Given the size of the dataset, we considered this approach more reliable than any crossvalidation approach.

@@

### Temperature
The temperature scaling function we utilized to define the relationship is the one defined by @Lloyd.1994l68 and it is derived from the original Arrhenius equation [Arrhenius.1889]:
$$
\xi_{temp(t,j)} = A_{k} \cdot e^{\frac{-Ea_j}{(T(t)+273.15)-T_0)}}
$$
The two parameters $A_{k}$ and $E_a$ are represented by Bayesian priors.  
For the temperature scaling we assumed that the activation energy term $E_a$ (which defines roughly the slope of the relationship) is driven by treatments ($j$) while the linear rescaling term $A_{k}$ is driven by the substrate at the single plot level ($k$). This because substrate (organic matter) amount is usually highly variable in soil even at small spatial scales, while its quality, which defines its temperature sensitivity, is assumed to follow the inputs quality which are relatively uniform over each treatments.  
The effect of the rescaling term $A_{k}$ is included in the function for temperature scaling to keep the convention of the original formulation of the equation, but it could be brought outside since it just a linear interaction.  
The prior distribution for the term $A$ was assumed as $\sim \mathcal{N}(300, 100)$ and for the term $E_a$ was assumed as $\sim \mathcal{N}(398.5, 50)$, utilizing as mean values the values  in @Lloyd.1994l68 and setting up conservatives standard deviations.

### Seasonality
In order to correct for eventual seasonality patters, we introduce also a seasonality scaling.
The seasonality scaling functions assume a sine function with period = 365 days, with both amplitude $ampl_j$ (the magnitude) and peak day $peak_j$ (when the function is at its only annual peak) varying as a function of treatments:
$$
\begin{aligned}
    \xi_{sin(t,k)} = ampl_k \cdot sin \left(\frac{2 \cdot \pi}{365}\right) \cdot day(t) \\
    +\left(\frac{2 \cdot \pi}{365}\right) \cdot (peak_k - 1) - \frac{\pi}{2})
\end{aligned}
$$
Both the terms $ampl_j$ and $peak_j$ are described by Bayesian priors. Variance is by treatment.
We assumed that the variability in seaslonalityis driven by treatments ($j$). 
The prior distribution for the term $ampl$ was assumed as $\sim \mathcal{N}(0, 0.5)$, so assuming as mean no seasonality effect, and for the term $peak$ was assumed as $\sim \mathcal{N}(175, 50)$, so assuming as mean for the peak the 24th of June (Midsummer day). Amplitude can be negative (this is just reverting the order of peaks and valleys of the sine function).

>>>>

Here, we hypothesized that:
(1) logging of trees will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated. In turn, addition of slash will provide labile organic matter to decompose increasing the respiration rates. 
(2) the variation in soil respiration would be mainly dependent on microclimatic variables (soil temperature and moisture), especially in the management treatments where vegetation was removed, since heterotrophic respiration will prevail over autotrophic respiration, and where the litter layer (slash) was removed
(3) the sensitivity of soil respiration to microclimatic conditions will also depend on the type of climate/forest, being the Mediterranean the less sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.
(4) the sensitivity of respiration to microclimate will decrease as the time passes since the disturbance because of new fresh C input from the growth of understory vegetation

To test these hypotheses, we have recorded data on soil respiration and microclimate every month in three replicated Quercus forests across Europe subjected to different forest management strategies during three years. We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic parametric model and compared to a non-parametric model (i.e., random forest).

@@

## Validation
In order to ensure a robust validation, all models have been validated with a split approach, keeping two completely independent data sets (80% of the points for training and 20% validation). Given the size of the dataset, we considered this approach more reliable than any cross-validation approach.

@@

### Temperature
The temperature scaling function we utilized to define the relationship is the one defined by @Lloyd.1994l68 and it is derived from the original Arrhenius equation [Arrhenius.1889]:
$$
\xi_{temp(t,j)} = A_{k} \cdot e^{\frac{-Ea_j}{(T(t)+273.15)-T_0)}}
$$
The two parameters $A_{k}$ and $E_a$ are represented by Bayesian priors.  
For the temperature scaling we assumed that the activation energy term $E_a$ (which defines roughly the slope of the relationship) is driven by treatments ($j$) while the linear rescaling term $A_{k}$ is driven by the substrate at the single plot level ($k$). This is because substrate (organic matter) amount is usually highly variable in soil even at small spatial scales, while its quality, which defines its temperature sensitivity, is assumed to follow the inputs quality which are relatively uniform over each treatments.  
The effect of the rescaling term $A_{k}$ is included in the function for temperature scaling to keep the convention of the original formulation of the equation, but it could be brought outside since it just a linear interaction.  
The prior distribution for the term $A$ was assumed as $\sim \mathcal{N}(300, 100)$ and for the term $E_a$ was assumed as $\sim \mathcal{N}(398.5, 50)$, utilizing as mean values the values  in @Lloyd.1994l68 and setting up conservatives standard deviations.

### Seasonality
In order to correct for eventual seasonality patters, we introduce also a seasonality scaling.
The seasonality scaling functions assume a sine function with period = 365 days, with both amplitude $ampl_j$ (the magnitude) and peak day $peak_j$ (when the function is at its only annual peak) varying as a function of treatments:
$$
\begin{aligned}
    \xi_{sin(t,k)} = ampl_k \cdot sin \left(\frac{2 \cdot \pi}{365}\right) \cdot day(t) \\
    +\left(\frac{2 \cdot \pi}{365}\right) \cdot (peak_k - 1) - \frac{\pi}{2})
\end{aligned}
$$
Both the terms $ampl_j$ and $peak_j$ are described by Bayesian priors. Variance is by treatment.
We assumed that the variability in seasonality is driven by treatments ($j$). 
The prior distribution for the term $ampl$ was assumed as $\sim \mathcal{N}(0, 0.5)$, so assuming as mean no seasonality effect, and for the term $peak$ was assumed as $\sim \mathcal{N}(175, 50)$, so assuming as mean for the peak the 24th of June (Midsummer day). Amplitude can be negative (this is just reverting the order of peaks and valleys of the sine function).
